Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-02-04 paper.qmd
Updated Figure 3

sc_ltr_pos <- c(0.08, 0.9)
# sc_xlims <- c(-0.5, 1.47)
# sc_ylims <- c(-0.32, 2.1)
sc_xlims <- c(-0.27, 1.2)
sc_ylims <- c(-0.2, 1.1)

@@

hex_grid_coloured_two_curvy1 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts, 
    aes(x = x, y = y,
      group = hex_poly_id, 
      fill = std_counts), color = "grey70", linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
             aes(x = tSNE1, y = tSNE2),
             alpha = 0.3,
             size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "E") +
  interior_annotation("b", sc_ltr_pos) + 
  xlim(c(-0.3, 1.2)) + ylim(c(-0.17, 1.1))

@@

@@

two_curvy_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.12, 0.12)) + 
  ylim(c(-0.12, 0.12)) +
  interior_annotation("a", c(0.08, 0.93), cex = 2) +
  theme(
    legend.position = "none"
  )

@@

## Second projection
projection <- cbind(
    c(-0.03372,0.07726,0.03948,-0.00096),
    c(0.07080,0.00316,-0.00382,0.05069))

@@

axes_obj <- gen_axes(  
  proj = projection * 6,
  limits = 0.1,
  axis_pos_x = -0.07,
  axis_pos_y = -0.07,
  axis_labels = names(scaled_two_curvy_data),
  threshold = 0.007)

two_curvy_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.09, 0.09)) + 
  ylim(c(-0.09, 0.09)) +
  interior_annotation("b", c(0.08, 0.93), cex = 2) +
  theme(
    legend.position = "none"
  )
```

>>>>

sc_ltr_pos <- c(0.08, 0.9)
# sc_xlims <- c(-0.5, 1.47)
# sc_ylims <- c(-0.32, 2.1)
sc_xlims <- c(-0.3, 1.2)
sc_ylims <- c(-0.17, 1.2)

@@

hex_grid_coloured_two_curvy1 <- ggplot() + 
  geom_polygon(
    data = hex_grid_with_counts, 
    aes(x = x, y = y,
      group = hex_poly_id, 
      fill = std_counts), color = "grey70", linewidth=0.2) +
  geom_point(data = tsne_two_curvy_scaled,
             aes(x = tSNE1, y = tSNE2),
             alpha = 0.3,
             size = 0.5) +
  scale_fill_viridis_c(direction = -1, 
    na.value = "#ffffff", option = "E") +
  interior_annotation("b", sc_ltr_pos) + 
  xlim(sc_xlims) + ylim(sc_ylims)

@@

## First projection
projection <- cbind(
  c(0.37847,-0.06517,0.04231,0.01655),
  c(-0.06466,-0.27052,0.22131,-0.15235))

@@

axes_obj <- gen_axes(  
  proj = projection * 2,
  limits = 0.6,
  axis_pos_x = -0.4,
  axis_pos_y = -0.4,
  axis_labels = names(scaled_two_curvy_data),
  threshold = 0.04)

two_curvy_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.52, 0.52)) + 
  ylim(c(-0.52, 0.52)) +
  interior_annotation("a", c(0.08, 0.93), cex = 2) +
  theme(
    legend.position = "none"
  )

@@

## Second projection
projection <- cbind(
  c(-0.38281,-0.04870,0.00619,-0.02451),
  c(-0.01303,0.00634,-0.37421,0.09648))

@@

axes_obj <- gen_axes(  
  proj = projection * 6,
  limits = 0.5,
  axis_pos_x = -0.35,
  axis_pos_y = -0.35,
  axis_labels = names(scaled_two_curvy_data),
  threshold = 0)

two_curvy_proj_umap_model2 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#2b8cbe") +
  geom_point(
    size = 0.5,
    alpha = 0.5,
    color = "#636363") +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2), 
    label=rownames(axes),
    colour="grey50",
    size = 4) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.4, 0.4)) + 
  ylim(c(-0.4, 0.4)) +
  interior_annotation("b", c(0.08, 0.93), cex = 2) +
  theme(
    legend.position = "none"
  )
```
